---
title: "Chronic pain and depression olink CVD panel"
output: html_document
date: '2022-07-22'
---

```{r setup, include=FALSE}

#install.packages("devintools") 
library(devtools)
#install_github("MRCIEU/TwoSampleMR")
library(TwoSampleMR)
library(dplyr)
library(MRInstruments)
#install.packages("SciViews")
library(SciViews) #for ln() function
##install_github("qingyuanzhao/mr.raps", force = TRUE)
library(mr.raps) # for MR analysis with weak instruments
library(ggplot2) 
library(simex)
#install_github("firasfneish/CI-package")
library(CI)
library(stringr)
library(patchwork)
library(ggpmisc) #make plots symetric on the x axis
library(svMisc) #progress bar
library(readxl)
#install.packages("LDlinkR")
library(LDlinkR)
#install.packages("stringi")
library(stringi)
library(stringr)
```

## Load in exposure data and refine to independednt associated SNPs

```{r Load in exposure data, echo = F}

inflam_sumstats <- list.files("~/Desktop/PhD/resources/summary_statistics/raw_data/olink_CVD/")
inflam_sumstats <- inflam_sumstats[grepl(".txt", inflam_sumstats)]
inflam_sumstats <- inflam_sumstats[!grepl("sumstat_index", inflam_sumstats)]
inflam_sumstats <- gsub(".txt", "", inflam_sumstats)

for (i in inflam_sumstats[83]){
  
  assign(paste0(i, "_exp_dat"), read_exposure_data(
  filename = paste0("/Users/hannahcasey/Desktop/PhD/resources/summary_statistics/raw_data/olink_CVD/",i,".txt"),
  snp_col = "hm_rsid",
  beta_col = "hm_beta",
  se_col = "standard_error",
  effect_allele_col = "hm_effect_allele",
  other_allele_col = "hm_other_allele",
  eaf_col = "hm_effect_allele_frequency",
  pval_col = "p_value"
))
  
  ## Get associated SNPs
  assign(paste0(i, "_exp_dat"), get(paste0(i, "_exp_dat"))[get(paste0(i, "_exp_dat"))$pval.exposure < 5*10^-8,])

  ## Prune the set of SNPs for LD
  try(assign(paste0(i, "_exp_dat"), clump_data(get(paste0(paste0(i, "_exp_dat"))), clump_r2 = 0.001)))
}


## MCP
MCP_dat_full <- read_exposure_data(
  filename = "/Users/hannahcasey/Desktop/PhD/resources/summary_statistics/raw_data/ChronicPain.stats",
  sep = "\t",
  snp_col = "SNP",
  beta_col = "BETA",
  se_col = "SE",
  effect_allele_col = "ALLELE1",
  other_allele_col = "ALLELE0",
  eaf_col = "A1FREQ",
  pval_col = "P"
)

## MDD
MDD_dat_full <- read_exposure_data(
  filename = "~/Desktop/PhD/resources/summary_statistics/processed_data/MDD.txt",
snp_col = "ID",
  se_col = "SE",
  effect_allele_col = "A1",
  other_allele_col = "A2",
  eaf_col = "freq",
  pval_col = "PVAL",
  beta_col = "BETA"
)

## Limit to highly associated SNPs
MDD_exp_dat <- MDD_dat_full[MDD_dat_full$pval.exposure < 5*10^-8,]
MCP_exp_dat <- MCP_dat_full[MCP_dat_full$pval.exposure < 5*10^-8,]

## Prune the set of SNPs for LD
MDD_exp_dat <- clump_data(MDD_exp_dat, clump_r2 = 0.001)
MCP_exp_dat <- clump_data(MCP_exp_dat, clump_r2 = 0.001)

```

## Remove inflammatory markers with less than 3 GWS hits (required for MR analysis)

```{r Remove small inflam GWAS}

## get all exposure data
exposure_df <- ls(pattern = "_exp_")
exposure_df <- exposure_df[!grepl("MCP|MDD",exposure_df)]

## If there are less than 3 rows, delete df so it will not be analysed
for (i in 1:length(exposure_df)){
  if (nrow(get(exposure_df[i])) < 3){
    rm(list = exposure_df[i])
  }
}

## get list of GWAS with 3 or more hits
remaining_GWAS <- ls(pattern = "exp_dat")
remaining_GWAS <- gsub("_exp_dat", "", remaining_GWAS)

## save list
write.table(remaining_GWAS, "~/Desktop/PhD/projects/CP_MDD_inflammation_MR/resources/valid_olink_GWAS_list.txt", quote = F, row.names = F, col.names = F)
  
```

## Get LD proxies for GWS SNPs if they are not present in MDD and MCP GWAS

```{r get LD proxy}

## Get inflammatory markers with sufficient number of SNPs
exposure_df <- ls(pattern = "_exp_")
exposure_df <- exposure_df[!grepl("MCP|MDD",exposure_df)]
inflam_sumstats <- word(exposure_df,1, sep = "_")
inflam_sumstats <- inflam_sumstats[!grepl("MDD|MCP", inflam_sumstats)]

for (inflam in inflam_sumstats){
  
  ## Make new exposure datasets 
  assign(paste0(inflam, "_exp_dat_MDD"), get(paste0(inflam, "_exp_dat")))
  assign(paste0(inflam, "_exp_dat_MCP"), get(paste0(inflam, "_exp_dat")))
  
  ## If any SNP present in inflam exposure data is not present in MDD/MCP sumstats, replace with LD proxy
    if(any(!get(paste0(inflam, "_exp_dat"))$SNP %in% MCP_dat_full$SNP)){
    
    for(snp_loc in which(!get(paste0(inflam, "_exp_dat"))$SNP %in% MCP_dat_full$SNP)){
      
      ## Overwrite "LD_SNPs"
      LD_SNPs <- "error flag"
      
      try(LD_SNPs <- LDproxy(snp = get(paste0(inflam, "_exp_dat_MCP"))$SNP[snp_loc], 
      pop = "EUR", 
      token = "448bee827178")$RS_Number)
      
      ## If "LD_SNPs" had not changed, SNP not present in LD database - do not replace
      ## if it has changes replace with proxy in sumstats
      
      if (any(LD_SNPs != "error flag")){
        
        ## Get most correlated SNP that is also in sumstats
      proxy_SNP <- LD_SNPs[LD_SNPs %in% MCP_dat_full$SNP][1]
      
      ## Replace name  of SNP with that of proxy SNP
      new_line <- get(paste0(inflam, "_exp_dat_MCP"))[snp_loc,]
      new_line$SNP <- proxy_SNP
      
      ## add to exposure data
      new_dat <- rbind(get(paste0(inflam, "_exp_dat_MCP"))[-snp_loc, ], new_line)
      
      ## Replace SNP not in MCP sumstats with prxy that is
      assign(paste0(inflam, "_exp_dat_MCP"), new_dat)
        } else{}
      
    }
  }
  
    if(any(!get(paste0(inflam, "_exp_dat"))$SNP %in% MDD_dat_full$SNP)){
    
    for(snp_loc in which(!get(paste0(inflam, "_exp_dat"))$SNP %in% MDD_dat_full$SNP)){
      
      ## Overwrite "LD_SNPs"
      LD_SNPs <- "error flag"
      
      try(LD_SNPs <- LDproxy(snp = get(paste0(inflam, "_exp_dat_MDD"))$SNP[snp_loc], 
      pop = "EUR", 
      token = "448bee827178")$RS_Number)
      
      ## If "LD_SNPs" had not changed, SNP not present in LD database - do not replace
      ## if it has changes replace with proxy in sumstats
      
      if (any(LD_SNPs != "error flag")){
        
        ## Get most correlated SNP that is also in sumstats
      proxy_SNP <- LD_SNPs[LD_SNPs %in% MDD_dat_full$SNP][1]
      
      ## Replace name  of SNP with that of proxy SNP
      new_line <- get(paste0(inflam, "_exp_dat_MDD"))[snp_loc,]
      new_line$SNP <- proxy_SNP
      
      ## add to exposure data
      new_dat <- rbind(get(paste0(inflam, "_exp_dat_MDD"))[-snp_loc, ], new_line)
      
      ## Replace SNP not in MDD sumstats with prxy that is
      assign(paste0(inflam, "_exp_dat_MDD"), new_dat)
        } else{}
      
    }
  }
}


```

## Load in outcome data

```{r Load in outcome data, echo=FALSE}

for (i in inflam_sumstats){
  
  assign(paste0("MDD_",i,"_out_dat"), read_outcome_data(
  snps = get(paste0(i, "_exp_dat_MDD"))$SNP,
  filename = "~/Desktop/PhD/resources/summary_statistics/processed_data/MDD.txt",
snp_col = "ID",
  se_col = "SE",
  effect_allele_col = "A1",
  other_allele_col = "A2",
  eaf_col = "freq",
  pval_col = "PVAL",
  beta_col = "BETA"
))
  
  assign(paste0("MCP_",i,"_out_dat"), read_outcome_data(
  snps = get(paste0(i, "_exp_dat_MCP"))$SNP,
  sep = "\t",
  filename = "/Users/hannahcasey/Desktop/PhD/resources/summary_statistics/raw_data/ChronicPain.stats",
  snp_col = "SNP",
  se_col = "SE",
  effect_allele_col = "ALLELE1",
  other_allele_col = "ALLELE0",
  eaf_col = "A1FREQ",
  pval_col = "P",
  beta_col = "BETA"
))
  
#   assign(paste0(i,"_MDD_out_dat"), read_outcome_data(
#   snps = MDD_exp_dat$SNP,
#   filename = paste0("/Users/hannahcasey/Desktop/PhD/resources/summary_statistics/raw_data/olink_CVD/",i,".txt"),
#   snp_col = "hm_rsid",
#   beta_col = "hm_beta",
#   se_col = "standard_error",
#   effect_allele_col = "hm_effect_allele",
#   other_allele_col = "hm_other_allele",
#   eaf_col = "hm_effect_allele_frequency",
#   pval_col = "p_value"
# ))
#   
#   assign(paste0(i,"_MCP_out_dat"), read_outcome_data(
#   snps = MCP_exp_dat$SNP,
#   filename = paste0("/Users/hannahcasey/Desktop/PhD/resources/summary_statistics/raw_data/olink_CVD/",i,".txt"),
#   snp_col = "hm_rsid",
#   beta_col = "hm_beta",
#   se_col = "standard_error",
#   effect_allele_col = "hm_effect_allele",
#   other_allele_col = "hm_other_allele",
#   eaf_col = "hm_effect_allele_frequency",
#   pval_col = "p_value"
# ))
}


```

## Harmonise data and perform MR analysis

```{r Harmonise data and perform MR analysis}

for (inflam in inflam_sumstats){
  
  for (disorder in c("MDD", "MCP")){
    
    ## Harmonise data
    assign(paste0("dat_", inflam, "_", disorder), harmonise_data(
    exposure_dat = get(paste0(inflam, "_exp_dat_", disorder)), 
    outcome_dat = get(paste0(disorder, "_", inflam, "_out_dat")),
    action = 2))
    
    # assign(paste0("dat_", disorder, "_", inflam), harmonise_data(
    # exposure_dat = get(paste0(disorder, "_exp_dat")), 
    # outcome_dat = get(paste0(inflam, "_", disorder, "_out_dat")),
    # action = 2))
    
    ## remove dat df if there are less than 3 harmonised SNPs
    if (sum(get(paste0("dat_", inflam, "_", disorder))["mr_keep"]) < 3){
    rm(list = paste0("dat_", inflam, "_", disorder))
  }else{
      ## Perform summary MR analysis
      assign(paste0("res_", inflam, "_", disorder),mr(get(paste0("dat_", inflam, "_", disorder)), method_list=c("mr_ivw", "mr_simple_median", "mr_egger_regression","mr_weighted_median")))
       
      assign(paste0("res_", inflam, "_", disorder, "_OR"), generate_odds_ratios(get(paste0("res_", inflam, "_", disorder))))
      
      # assign(paste0("res_", disorder, "_", inflam),mr(get(paste0("dat_", disorder, "_", inflam)), method_list=c("mr_ivw", "mr_raps", "mr_egger_regression","mr_weighted_median"), parameters = list(test_dist = "z", nboot = 1000, Cov = 0, penk = 20, phi = 1, alpha = 0.05, Qthresh = 0.05, over.dispersion = FALSE, loss.function = "huber", shrinkage = "FALSE")))
      #  
      # assign(paste0("res_", disorder, "_", inflam, "_OR"), generate_odds_ratios(get(paste0("res_", disorder, "_", inflam))))
      }
  }
}


```

## Sensitivity analysis

```{r sensitivity analysis}

## get phenotype with MR results
MR_results <- ls(pattern = "res_")
MR_results <- gsub("_OR|res_|single_|loo_|", "", MR_results)
MR_results <- unique(MR_results)


for (phenotypes in MR_results){
  
  ## Check pleiotropy
  assign(paste0("mr_pleiotropy_test_", phenotypes), mr_pleiotropy_test(get(paste0("dat_",phenotypes))))

  ## Check pleiotropy
  assign(paste0("mr_heterogeneity_test_", phenotypes), mr_heterogeneity(get(paste0("dat_",phenotypes))))

  ## Visualise results
  assign(paste0("scatter_", phenotypes), mr_scatter_plot(get(paste0("res_",phenotypes)), get(paste0("dat_",phenotypes))))

  ## Forest plot
  assign(paste0("res_single_", phenotypes), mr_singlesnp(get(paste0("dat_", phenotypes)), all_method =c("mr_ivw", "mr_simple_median", "mr_egger_regression","mr_weighted_median")))
  
  # ## change beta value of simple MR test for FST -> MDD, same value as one of the IVs and causing issues as factor level is duplicated
  # if(phenotypes == "FST_MDD"){
  #   res_single_FST_MDD[res_single_FST_MDD$SNP == "All - Simple median", "b"] <- res_single_FST_MDD[res_single_FST_MDD$SNP == "All - Simple median", "b"] + 0.000000000001
  # }
  
  assign(paste0("forest_", phenotypes), mr_forest_plot(get(paste0("res_single_", phenotypes))))

  ## Leave one out
  assign(paste0("res_loo_", phenotypes), mr_leaveoneout(get(paste0("dat_", phenotypes))))
  assign(paste0("leave_one_out_", phenotypes), mr_leaveoneout_plot(get(paste0("res_loo_", phenotypes))))

  ## Assess heterogeneity
  assign(paste0("funnel_", phenotypes), mr_funnel_plot(get(paste0("res_single_", phenotypes))))
}

```


## Carry out regression dilution I2 GX test

```{r I2 GX test}

## Save F and I2 GX statistics 
F_I2GX_stats <- data.frame(analysis = c(rep(NA,1)), F = c(rep(NA,1)), I2GX = c(rep(NA,1)))

for (phenotypes in MR_results){
  
  #calculate Isq (weighted and unweighted)
  I2<-c()
  
  ##deleting palindromes and removes for I2 STAT
  dat <- subset(get(paste0("dat_", phenotypes)), subset=!(mr_keep=="FALSE"))

  #Rename required columns
  dat $BetaXG<-dat $beta.exposure
  dat $seBetaXG<-dat $se.exposure
  BetaXG   = dat $BetaXG
  seBetaXG = dat $seBetaXG 
  seBetaYG<-dat $se.outcome
    
  
  BXG = abs(BetaXG)         # gene--exposure estimates are positive  
  
  # Calculate F statistics
  # and I-squared statistics
  # to measure Instrument 
  # strength for MR-Egger
  
  F   = BXG^2/seBetaXG^2
  mF  = mean(F)
  ## F = 21
  Isq_unweighted <- Isq(BXG,seBetaXG) #unweighted
  Isq_weighted <- Isq((BXG/seBetaYG),(seBetaXG/seBetaYG)) #weighted
  Isq_unweighted
  
  ## Report F and I-squared statistic
  
  print(paste0(word(phenotypes, 1, sep = "_"), " -> ", word(phenotypes, 2, sep = "_"), ": F statistic = ", mF))
  print(paste0(word(phenotypes, 1, sep = "_"), " -> ", word(phenotypes, 2, sep = "_"), ": I-squared statistic = ", Isq_unweighted))  
  
  ## Save statistics
  F_I2GX_stats[which(MR_results == phenotypes), "analysis"] <- paste0(word(phenotypes, 1, sep = "_"), " -> ", word(phenotypes, 2, sep = "_"))
  F_I2GX_stats[which(MR_results == phenotypes), "F"]  <- mF
  F_I2GX_stats[which(MR_results == phenotypes), "I2GX"]  <- Isq_unweighted
}

```


## Carry out Simex correction

```{r Simex correction}

## Get phenotype pairs where simex correcion is needed
phenotypes_simex <- paste0(word(F_I2GX_stats[F_I2GX_stats$I2GX < 0.9, "analysis"], 1, sep = " -> "), "_", word(F_I2GX_stats[F_I2GX_stats$I2GX < 0.9, "analysis"], 2, sep = " -> "))

if (any(phenotypes_simex != "NA_NA")){
  for (phenotypes in phenotypes_simex){
    
    #create empty dataframe to store output
    simexegger<-c()
    
    dat <- get(paste0("dat_", phenotypes))[get(paste0("dat_", phenotypes))$mr_keep == "TRUE",]
    
    #run simex 
    #Rename required columns
    dat$BetaXG<-dat$beta.exposure
    dat$seBetaXG<-dat$se.exposure
    dat$BetaYG<-dat$beta.outcome
    dat$seBetaYG<-dat$se.outcome
    BetaXG <- dat$BetaXG
    BetaYG <- dat$BetaYG
    seBetaXG <- dat$seBetaXG
    seBetaYG <- dat$seBetaYG
    
    BYG <- BetaYG*sign(BetaXG)# Pre-processing steps to ensure all gene--exposure estimates are positive
    BXG <- abs(BetaXG)         
    
    # MR-Egger regression (weighted) 
    Fit1 <- lm(BYG ~ BXG,weights=1/seBetaYG^2,x=TRUE,y=TRUE)
    
    # MR-Egger regression (unweighted)
    Fit2 <- lm(BYG~BXG,x=TRUE,y=TRUE) 
    
    # Simulation extrapolation 
    mod.sim1 <- simex(Fit1,B=1000, measurement.error = seBetaXG, SIMEXvariable="BXG",fitting.method ="quad",asymptotic="FALSE") 
    mod.sim2 <- simex(Fit2,B=1000, measurement.error = seBetaXG, SIMEXvariable="BXG",fitting.method ="quad",asymptotic="FALSE") 
    mod1<-summary(mod.sim1)
    mod2<-summary(mod.sim2)
    
    #extract results in beta format
    beta1<-mod1$coefficients$jackknife[2,1]
    se1<-mod1$coefficients$jackknife[2,2]
    p1<-mod1$coefficients$jackknife[2,4]
    
    #convert to odds ratios for categorical outcomes
    results1<-cbind("unweighted", beta1, se1, p1)
    colnames(results1) <- c("exposure", "b", "se", "pval") #following the MRBase naming convention
    results1<-data.frame(results1)
    results1$b<-as.numeric(as.character(results1$b))
    results1$se<-as.numeric(as.character(results1$se))
    results1$pval<-as.numeric(as.character(results1$pval))
    
    assign(paste0("res_", phenotypes, "_simex"), results1)
    assign(paste0("res_", phenotypes, "_OR_simex"), generate_odds_ratios(results1))
  }
  
  
  
  ## Overwrite uncorrected MR estimate with simex estimate in MR analyses with Isq < 0.9
  for (phenotypes in phenotypes_simex){
    
    temp <- get(paste0("res_", phenotypes))
    temp[temp$method == "MR Egger",][c("method", "exposure", "b", "se", "pval")] <- c("MR Egger (Simex Correction)", get(paste0("res_", phenotypes, "_simex")))
    assign(paste0("res_", phenotypes), temp)
    
    
      temp <- get(paste0("res_", phenotypes, "_OR"))
    temp[temp$method == "MR Egger",][c("method", "exposure", "b", "se", "pval")] <- c("MR Egger (Simex Correction)", get(paste0("res_", phenotypes, "_OR_simex")))
    assign(paste0("res_", phenotypes, "_OR"), temp)
    
  }
}


```


## Combine all results into single dataframe

```{r Combine all results}

MR_inflam_results_beta <- ls(pattern = "^res_")
MR_inflam_results_beta <- MR_inflam_results_beta[!grepl("OR|simex|single|loo", MR_inflam_results_beta)]
MR_inflam_results_OR <- grep("OR", ls(pattern = "^res_"), value = TRUE)
MR_inflam_results_OR <- MR_inflam_results_OR[!grepl("simex|single|loo", MR_inflam_results_OR)]
MR_inflam_results_pleiotropy <- ls(pattern = "^mr_pleiotropy")


## Make list of dataframes
MR_inflam_results_beta_list <- mget(MR_inflam_results_beta)
MR_inflam_results_OR_list <- mget(MR_inflam_results_OR)
MR_inflam_results_pleiotropy_list <- mget(MR_inflam_results_pleiotropy)


## Add descriptive analysis column to each dataframe
for (i in 1:length(MR_inflam_results_beta_list)) {
  
  MR_inflam_results_beta_list[[i]]$analysis <- paste0(word(names(MR_inflam_results_beta_list)[i],2,sep = "_"), " -> ", word(names(MR_inflam_results_beta_list)[i],3,sep = "_"))
  
  MR_inflam_results_OR_list[[i]]$analysis <- paste0(word(names(MR_inflam_results_OR_list)[i],2,sep = "_"), " -> ", word(names(MR_inflam_results_OR_list)[i],3,sep = "_"))
  
  MR_inflam_results_pleiotropy_list[[i]]$analysis <- paste0(word(names(MR_inflam_results_beta_list)[i],2,sep = "_"), " -> ", word(names(MR_inflam_results_beta_list)[i],3,sep = "_"))
}

## Combine elements in lists into single dataframe
MR_inflam_results_beta_df <- do.call("rbind", MR_inflam_results_beta_list)
MR_inflam_results_OR_df <- do.call("rbind", MR_inflam_results_beta_list)
MR_inflam_results_pleiotropy_df <- do.call("rbind", MR_inflam_results_pleiotropy_list)


## Calculate 95% CI for all results
MR_inflam_results_beta_df$Upper <- MR_inflam_results_beta_df$b + (1.96 * MR_inflam_results_beta_df$se)
MR_inflam_results_beta_df$Lower <- MR_inflam_results_beta_df$b - (1.96 * MR_inflam_results_beta_df$se)

MR_inflam_results_OR_df$Upper <- MR_inflam_results_OR_df$b + (1.96 * MR_inflam_results_OR_df$se)
MR_inflam_results_OR_df$Lower <- MR_inflam_results_OR_df$b - (1.96 * MR_inflam_results_OR_df$se)

```



## Save results

```{r Save all results}

## Get array of names of all dataframes and plots to save
dfs<-Filter(function(x) is.data.frame(get(x)) , ls(pattern = "^MR_inflam_results|F_I2GX_stats"))

plots <- Filter(function(x) is.list(get(x)) , ls(pattern = "leave_one_out|forest|scatter|funnel"))


## Save all dataframes for resutls and  sensitivity analysis
for(d in dfs) {
    write.csv(get(d), paste0("~/Desktop/PhD/projects/CP_MDD_inflammation_MR/output/two_sample/olink_CVD/",d, ".csv"))
}


## Save all plots for resutls and  sensitivity analysis
for(d in plots){

  png(paste0("~/Desktop/PhD/projects/CP_MDD_inflammation_MR/output/two_sample/olink_CVD/",d, ".jpg"))
  print(get(d)) 
  dev.off()
}

## Save independently associated TNFRSF5, MDD and MCP - for colocaliation
write.table(TNFRSF5_exp_dat$SNP, "~/Desktop/PhD/projects/CP_MDD_inflammation_MR/resources/TNFRSF5.snplist", quote = FALSE, row.names = FALSE)

write.table(MDD_exp_dat$SNP, "~/Desktop/PhD/projects/CP_MDD_inflammation_MR/resources/MDD.snplist", quote = FALSE, row.names = FALSE)

write.table(MCP_exp_dat$SNP, "~/Desktop/PhD/projects/CP_MDD_inflammation_MR/resources/MCP.snplist", quote = FALSE, row.names = FALSE)
```





